---
title: "Hw01"
author: "Maliha Tariq"
date: "2/16/2018"
output:
  html_document: null
  keep_md: yes
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(plotly)
library(forcats)
library(tidyverse)
library(ggrepel)
library(DT)
```

```{r}
library(readr)
winter <- read_csv("~/Documents/GitHub/Spring2018/Tariq_Maliha/hw01/winter.csv")
dictionary <- read_csv("~/Documents/GitHub/Spring2018/Tariq_Maliha/hw01/dictionary.csv")
```

```{r}
dictionary$CountryName <-dictionary$Country
dictionary$Country <-dictionary$Code
dictionary$Code <-NULL
```

```{r}
winter$Country <- recode(winter$Country, FRG = "GER")
winter$Country <- recode(winter$Country, GDR = "GER")
winter$Country <- recode(winter$Country, EUA = "GER")
winter$Country <- recode(winter$Country, EUN = "RUS")
winter$Country <- recode(winter$Country, TCH = "CZE")
```

```{r}
merged <- merge(dictionary, winter, by = c('Country'))
```

##1. Medal Counts Over Time

Graph 1- Total Medals for Top 10 Countries

```{r}
merged %>%
  group_by(CountryName) %>%
  summarise(Total= n()) %>%
  mutate(CountryName = fct_reorder(CountryName, Total, .desc = TRUE)) %>%
  top_n(10, Total) %>%
  ggplot(aes(x = CountryName, y = Total, fill=CountryName)) + 
  geom_bar(stat = "identity") + 
  theme_classic() +
  theme(legend.position = "") +
  ggtitle("Total Medals for Top 10 Countries") +
  xlab("Country") +
  ylab("Number of Medals") +
  theme_classic() +
  theme (plot.title = element_text(hjust = .5))
```

Graph 2- Comparison of Top 10 Countries Total


```{r}
merged %>%
  group_by(CountryName) %>%
  mutate(Total = n()) %>%
  distinct(CountryName, .keep_all=TRUE) %>%
  filter(Country == "USA" | Country == "GER" | Country == "CAN"| Country == "NOR" | Country == "FIN" | Country == "RUS"| Country == "SUI" | Country == "SWE"| Country == "AUT" | Country == "CZE") %>%
  ggplot(merged, mapping = aes(CountryName, Total, label = unique(CountryName))) +
  geom_text_repel() +
  geom_point(aes(colour = factor(CountryName))) +
  theme_classic(base_size = 15) +
  theme (axis.title.x=element_blank(),axis.text.x=element_blank()) +
  theme(legend.position = "") +
  ggtitle("Total Medals for Top 10 Countries") +
  theme (plot.title = element_text(hjust = .5)) +
  ylab("Number of Medals")

```

Graph 3- Medals Over Time

```{r name of chunk}
time_plot <- merged %>%
  group_by(Country, Year) %>%
  mutate(Total = n()) %>%
  distinct(CountryName, .keep_all=TRUE) %>%
  filter(Country == "USA" | Country == "GER" | Country == "CAN"| Country == "NOR" | Country == "FIN" | Country == "RUS"| Country == "SUI" | Country == "SWE"| Country == "AUT" | Country == "CZE") %>%
  ggplot(merged, mapping = aes(x = Year, y = Total)) +
  geom_line(aes(color=Country, group=Country)) +
  theme(legend.position = "") +
  ggtitle("Total Medals for Top 10 Countries Over Time") +
  ylab("Number of Medals") +
  theme_classic() +
  theme (plot.title = element_text(hjust = .5))


time_plot
```

Graph 4- Type of Medals Won by Top Performing Countries

```{r}
order <- c("USA", "GER", "CAN", "NOR", "FIN", "SWE", "RUS", "SUI", "AUT", "CZE")
m_order <- c("Gold", "Silver", "Bronze")
merged %>%
  group_by(Country, Medal) %>%
  summarize(Total = n()) %>%
  filter(Country == "USA" | Country == "GER" | Country == "CAN"| Country == "NOR" | Country == "FIN" | Country == "RUS"| Country == "SUI" | Country == "SWE"| Country == "AUT" | Country == "CZE") %>%
  ungroup() %>%
  mutate(Country = factor(Country, levels = order)) %>%
  mutate(Medal = factor(Medal, levels = m_order)) %>%
  ggplot(aes(x = Country, y = Total, fill = Medal)) +
  geom_bar(stat="identity") +
  scale_fill_manual(values=c("Gold", "light gray","brown")) +
  scale_color_gradient() +
  theme_classic() +
  ggtitle("Type of Medal by Country") +
  theme (plot.title = element_text(hjust = 0.5)) +
  xlab("Country") + ylab("Number of Medals")
```

Graph 1 summarizes the total number of medals for each of the top 10 performing countries throughout the Winter Olympics. Graph 2 represents the same information but using points instead of bars. Graph 3 shows the top ten countries and the total number of medals they have received over time from 1924 to 2014. Finally, Graph 4 illustrates types of medals each of the top 10 countries have received. Of the four graphs represented, the first graph may be the most intuitive for readers, since bar graphs allow for a quick understanding and comparison of values. However, Graph 3 illustrates interesing information in trends over time. Additionally, Graph 4 helps illustrate the ratio of Gold medals to Silver Medals and Bronze Medals; this is interesting and useful information for comparisions as well.

##2. Medal Counts Adjusted by Population and GDP

#Who's Taking Home the Most Gold?

(a)-Unadjusted Rankings

```{r}
unadjusted_Gold <-
order3 <- c("CAN", "GER", "USA", "NOR", "RUS", "SWE", "AUT", "SUI", "FIN", "ITA")
merged %>%
  filter(Medal == "Gold") %>%
  group_by(Country) %>%
  summarize(Count = n()) %>%
  top_n(10, Count) %>%
  arrange(desc(Count)) %>%
  mutate(Country = factor(Country, levels = order3)) %>%
  ggplot(aes(x = Country, y= Count, fill = Country)) +
  geom_bar(stat = "identity") +
  theme_gray() +
  ggtitle("Who's Taking Home the Most Gold?") +
  xlab("Country") + ylab("Number of Gold Medals") +
  theme(legend.position = "") +
  theme_classic() +
  theme (plot.title = element_text(hjust = 0.5))

unadjusted_Gold
```

According to the unadjusted data, Canada has the most gold medals, followed by Germany, then the United States, Norway, Russia, Sweden, Austria, Switzerland, Finland, and Italy. We can see that this order is different than the total medal count rankings.

(b)-Adjusted with Respect to Population

```{r}
merged <- 
  transform(merged, new_pop = merged$Population/1000000) %>%
  transform(merged, adjusted_GDP = GDP.per.Capita /10000)
```


```{r}
data_pop <- 
order <- c("LIE", "NOR", "SWE", "FIN", "AUT", "SUI", "CAN", "EST", "CZE", "GER")
merged %>%
  transform(merged, new_pop = merged$Population/1000000) %>%
  filter(Medal == "Gold") %>%
  group_by(Country) %>%
  mutate(Count = n()) %>%
  distinct(Country, .keep_all=TRUE) %>%
  mutate(Count / new_pop) %>%
  transform(medals_pop = Count / new_pop) %>%
  top_n(10, medals_pop) %>%
  mutate(Country = factor(Country, levels = order)) %>%
  arrange(desc(medals_pop)) %>%
  ggplot(aes(x = Country, y = medals_pop, fill = Country)) + 
  geom_bar(stat = "identity") +
  theme_classic() +
  ggtitle("Gold Medal Wins with Respect to Population") +
  theme (plot.title = element_text(hjust = 0.5)) +
  xlab("Country") + ylab("Medals per Million people") +
  theme(legend.position = "")
```

When adjusting for population, Liechtenstein, Norway and Sweden ranked in the top 3, illustrating the relationship between their population and the number of gold medals they received. It makes sense that a country as small as Liechtenstein ranks at the top since their population is very small. Its also interesting to note that countries like the US and Russia no longer rank in the top 10 since they have relatively larger populations, even though they have more medals than a few of the countries included above.

(c)Adjusted with Respect to GDP

```{r}
order <- c("RUS", "CAN", "GER", "USA", "SWE", "UKR", "NOR", "CHN", "ITA", "KOR")
merged %>%
  transform(merged, adjusted_GDP = GDP.per.Capita /10000) %>%
  filter(Medal == "Gold") %>%
  group_by(Country) %>%
  mutate(Count = n()) %>%
  distinct(Country, .keep_all=TRUE) %>%
  transform(medals_gdp = Count/ adjusted_GDP) %>%
  mutate(Count / adjusted_GDP) %>%
  top_n(10, medals_gdp) %>%
  mutate(Country = factor(Country, levels = order)) %>%
  arrange(desc(medals_gdp)) %>%
  ggplot(aes(x = Country, y = medals_gdp, fill = Country)) + 
  geom_bar(stat = "identity") +
  theme_classic() +
  ggtitle("Gold Medal Wins with Respect to GDP") +
  theme (plot.title = element_text(hjust = 0.5)) +
  xlab("Country") + ylab(" Medals per GDP per Capita (per 10,000)") +
  theme(legend.position = "") 
```

When factoring in GDP, Russia, Canda and Germany rank in the Top 3. However, this figure is representing the ratio of Medals over GDP, so the higher ranking countries do not necessarily have higher GDP values; they have a higher medal to GDP per capita (divided by a factor of 10,000).

##3. Host Country Advantage

```{r}
library(rvest)
library(stringr)
wiki_hosts <- read_html("https://en.wikipedia.org/wiki/Winter_Olympic_Games")
hosts <- html_table(html_nodes(wiki_hosts, "table")[[5]], fill=TRUE)
hosts <- hosts[-1,1:3]
hosts$city <- str_split_fixed(hosts$Host, n=2, ",")[,1]
hosts$country <- str_split_fixed(hosts$Host, n=2, ",")[,2]

names(hosts)[names(hosts) == "city"] <- "City"
merged <- merge(hosts, merged, by = 'City')
names(merged)[names(merged) == "country"] <- "Hostcountry" 

merged$Hostcountry <- trimws(merged$Hostcountry, which = "left")
merged$Hostbin <- ifelse(merged$Hostcountry == merged$CountryName, 1, 0)
```

```{r}
  merged %>%
  group_by(CountryName, Year.y) %>% 
  mutate(Count = n()) %>%
  distinct(CountryName, .keep_all=TRUE) %>%
  ggplot(merged, mapping = aes(x = as.numeric(Year.y), y = Count, color = as.factor(Hostbin))) +
  geom_jitter() +
  scale_x_continuous(breaks = seq(1924, 2014, by=9)) + 
  geom_smooth(method = 'lm', se = FALSE) +
  theme_classic() +
  ggtitle("Does Hosting Give You An Advantage?") +
  theme (plot.title = element_text(hjust = 0.5)) +
  xlab("Years") + ylab("Number of Medals") +
  scale_color_manual(values = c("orange", "turquoise"), labels =c ("Not Hosting", "Hosting")) +
  theme(legend.title = element_blank())
```

The graph above illstrates the difference between the number of medals won by countries who were hosting the Olympics versus those who were not. The orange dots represent the values for the countries who were not hosting during a given year, while the turqoise dots represent those that were hosting. It is interesting to compare the relative placement of those dots. The graph also inclues regression lines to show the difference between the two groups. Therefore, we can see that countries who host the Olympics have a slightly higher number of medal wins, on average, than those who are not hosting in a given year. 

##4- Country Success by Sport/Discipline/Event

```{r}
merged %>%
  group_by(CountryName, Sport, Gender) %>%
  count() %>% arrange(-n) %>% 
  filter(Sport == "Skiing") %>%
  head(16) %>% 
  ggplot(aes(x = CountryName, y = n, fill = Gender)) +
  geom_col()+
  theme_classic() +
  ggtitle("Male Vs Female Medal Wins for Skiing in the Top 10 Countries") +
  theme (plot.title = element_text(hjust = 0.5)) +
  xlab("Country") + ylab("Number of Medals")
  
```

The bar graph above illustrates the number of medals won for skiing in the top 10 ranking countries and further separates that data by gender for male versus female athletes. It is interesteing to see the differences by gender and the fact that France and Italy had all male winners, while Russia had all female winners.

##5- Most Successful Athletes

```{r}
merged %>%
  group_by(Athlete, Gender) %>%
  count() %>% arrange(-n) %>% 
  head(10) %>% 
  ggplot(aes(x = Athlete, y = n, fill = Gender)) + coord_flip() +
  geom_col()+
  theme_classic() +
  ggtitle("Top Ranking Athletes by Gender") +
  theme (plot.title = element_text(hjust = 0.5))
```

It is interesting to see that in the top 10 rankings for individual athletes, there are both five males and five males. Additionally, while the top 2 athletes are male, the third and fourth highest ranked athletes are female.

##6-Interactivity

```{r}
ggplotly(time_plot)
```

The first interactive plot illustrates the number of medals won by each top 10 country over time.

```{r}
f <- list(
  family = "Courier New, monospace",
  size = 18,
  color = "#7f7f7f"
)
x <- list(
  title = "Years",
  titlefont = f
)
y <- list(
  title = "Medals Won",
  titlefont = f
)
  merged %>%
  group_by(CountryName, Year.y) %>% 
  mutate(Count = n()) %>%
  distinct(CountryName, .keep_all=TRUE) %>%
  
  plot_ly(x = ~as.numeric(Year.y), y = ~Count, color = ~as.factor(Hostbin)) %>%
  add_markers(text = ~Country, hoverinfo = "text") %>%
    layout(xaxis = x, yaxis = y) %>%
      layout(showlegend = FALSE) %>%
      layout(title = "Hosts Vs. Non-Hosts")

```

The second interactive plot illustrates how host countries perform relative to non-host countries. The blue dot represent the host countries, while the green dots represent the non-host countries. You can hover over the dots to see the identity of the country that is hosting vs. not hosting.

##7- Data Table

```{r}
datatable(merged) %>%
  formatStyle('CountryName',  color = 'white', 
                backgroundColor = 'Green', fontWeight = 'bold',
              )
```

In this data table, I am highlighting the country column to allow users to better filter by their country of interest and view stats on the population, GDP, the years the country participated, etc.
